---
title: "Monte Carlo Integration Practice"
author: "Mikayla Mulgrew"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction
Monte Carlo (MC) simulations provide a means to model a problem and apply brute force computational power to achieve a solution - randomly simulate from a model until you get an answer. The best way to explain is to just run through a bunch of examples, so lets go!

# Integration
We will start with basic integration. Suppose we have an instance of a Normal distribution with a mean of 1 and a standard deviation of 2 then we want to find the _integral_ (area under the curve) from 1 to 3:
\[\int_1 ^3 \frac{1}{10\sqrt{2\pi}}e^{-\frac{(x-1)^2}{2\times 2^2}}\, dx\]


First, create a new R script in Rstudio. Next we define the number of samples we will obtain. Lets choose 100.

Then use the R function `rnorm` to simulate 100 numbers from a Normal distribution with mean 1 and standard deviation 2.

Estimate the integral between 1 and 3 by counting how many samples had a value in this range.

The `pnorm` gives the integral under the Normal distribution (in this case with mean 1 and standard deviation 2) from negative infinity up to the value specified by q.

# MC integration problem

> **Q1.** Try increasing the number of simulations and see how the accuracy improves?

> **Q2.** Can you draw a graph of number of MC samples vs accuracy?

```{r}

#define number of samples to obtain
n <- 100

#use rnorm to simulate 100 numbers from a normal distribition, mean=1 and sd=2
simulations <- rnorm(n, mean=1, sd=2)

#estimate integral between 1 and 3 by counting samples with a value in range
mc_S <- sum(simulations <=3 & simulations >= 1)/100
print(paste("Simulation with 100 samples: ", mc_S))

#The exact answer given using the cumulative distribution function pnorm in R is given by:
mc_exact = pnorm(q=3, mean=1, sd=2) - pnorm(q=1, mean=1, sd=2)
print(paste("Exact answer using CDF: ", mc_exact))

##Q1 - increasing sample size to increase accuracy
n2 <- 10000

simulations_2 <- rnorm(n2, mean=1, sd=2)
mc_S2 <- sum(simulations_2 <=3 & simulations_2 >=1)/n2
print(paste("Simulation with 10,000 samples: ", mc_S2))

#Q2 - Can you draw a graph of number of MC samples vs accuracy?

#create a vector to hold simulated value from each simulation
values <- c()

for (i in seq(from=100, to=100000, by=1000)){
  
  sims <- rnorm(i, mean=1, sd=2)
  mcS <- sum(sims <=3 & sims >=1)/i
  #add difference of mcexact to mcS to values vector
  difference <- mc_exact - mcS
  
  val <- as.character(i)
  values[val] <- difference
  
}

#plot barplot showing difference in simulate values vs exact value from CDF
barplot(values, main="Difference between Exact Value and MC Simulated Values", xlab="Simulation Sample Size", ylab="Difference")
```

Model Solution plots difference in mean and sd, try that. 
  
Example:  
  
  accuracy[i] = mean(mc_integral - mc_exact) # compute average difference between integral estimate and real value
  accuracy_sd[i] = sd(mc_integral - mc_exact) # compute sd difference between integral estimate and real value

}

print(accuracy)

print(accuracy_sd)

print(accuracy+accuracy_sd)

# load ggplot
library(ggplot2)

# create a data frame for plotting
df = data.frame(sample_sizes, accuracy, accuracy_sd)

print(df)

# use ggplot to plot lines for the mean accuracy and error bars using the std dev
ggplot(df, aes(x=sample_sizes, y=accuracy)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin=accuracy-accuracy_sd, ymax=accuracy+accuracy_sd), width=.2,
                position=position_dodge(0.05)) +
  ylab("Estimate-Exact") +
  xlab("Run")
  
  
  
  
# Approximating the Binomial Distribution
We flip a coin 10 times and we want to know the probability of getting more than 3 heads. This is a trivial problem using the Binomial distribution but suppose we have forgotten about this or never learned it in the first place.

Let's solve this problem with a Monte Carlo simulation. We will use the common trick of representing tails with 0 and heads with 1, then simulate 10 coin tosses 100 times and see how often that happens.

```{r}

#number of simulations to run
n_sims <- c(10, 100, 500, 1000, 2000, 5000)

#create a vector to hold outcome for each simulation, mean, sd
great_3 <- rep(0, length(n_sims))
s_mean_acc <- rep(0, length(n_sims))
s_sd_acc <- rep(0, length(n_sims))

#use for loop to run through each simulation size
for (s in 1:length(n_sims)){
  
  #get simulation size
  simsz <- n_sims[s]
  
  #vector to hold outcome of coin toss
  greater_3 <- rep(0, simsz)
  
  #toss coin for number of simulations
  for (x in 1:simsz){
    
    #flip a coin ten times
    flips <- sample(c(0,1), 10, replace=TRUE)
  
    #if more than 3 heads are flipped change value to 1 in great_3
    if (sum(flips) > 3){
      greater_3[x] <- 1
    }
    
  }
  
  #get mean and sd of simulated values and add to vector s_mean_acc and s_sd_acc
  
  s_mean_acc[s] <- mean(greater_3 - pbinom(3, 10, 0.5, lower.tail=FALSE))
  s_sd_acc[s] <- sd(greater_3 - pbinom(3, 10, 0.5, lower.tail=FALSE))
  
}

print(s_mean_acc)
print(s_sd_acc)

#true value
print( pbinom(3, 10, 0.5, lower.tail=FALSE) )

```

## Problem
> **Q1.** Try increasing the number of simulations and see how the accuracy improves?

> **Q2.** Can you plot how the accuracy varies as a function of the number of simulations? (hint: see the previous section)


```{r}

library(ggplot2)

#create data frame with simulation sizes and means
great_df <- data.frame(n_sims, s_mean_acc)

#plot mean accuracy with sd error bars
ggplot(great_df, aes(x=n_sims, y=s_mean_acc)) +geom_line() + geom_point() + geom_errorbar(aes(ymin=s_mean_acc-s_sd_acc, ymax=s_mean_acc+s_sd_acc)) + xlab("Simulation Size") + ylab("Estimate - Exact")

```
# Monte Carlo Expectations
Consider the following spinner. If the spinner is spun randomly then it has a probability 0.5 of landing on yellow and 0.25 of landing on red or blue respectively.

If the rules of the game are such that landing on 'yellow' you gain 1 point, 'red' you lose 1 point and 'blue' you gain 2 points. We can easily calculate the expected score.

Let X denote the random variable associated with the score of the spin then:

\[E[X]=(\frac{1}{2} \times 1)+(\frac{1}{4}\times-1)+(\frac{1}{4}\times2)=0.75\]

If we ask a more challenging question such as:

> After 20 spins what is the probability that you will have less then 0 points?"

How might we solve this?

Of course, there are methods to analytically solve this type of problem but by the time they are even explained we could have already written our simulation!

To solve this with a Monte Carlo simulation you need to sample from the Spinner 20 times, and return 1 if we are below 0 other wise we'll return 0. We will repeat this 10,000 times to see how often it happens!

## Functions
First, we are going to introduce the concept of a function. This is a piece of code which is encapsulated so then we can refer to it repeated via the name of the function rather than repeatedly writing those lines of code. The function we will write will simulate one game as indicated above and return whether the number of points is less than zero.
## Simulation
Now we can use this function in a loop to play the game 100 times.

We can then compute the probability that, after twenty spins, we will have less than zero points

## Problems
> **Q1.** Modify your code to allow you to calculate the expected number of points after 20 spins.
> **Q2.** Simulate a game in which you have a maximum of 20 spins but you go “bust” once you hit a negative score and take this into account when you compute the expected end of game score.

```{r}

game <- function(){
  
  #random sample to model spinning wheel
  #Y, 1pt, prob=0.5
  #R, -1pt, prob=0.25
  #B, 2pt, prob=0.25
  
  results <- sample(c(1,-1,2), 20, replace=TRUE, prob=c(0.5,0.25,0.25))
  
  #get cumulative outcomes for the 20 spins 
  cumulative <- cumsum(results)
  
  #check if any are 0 or less than zero
  for (i in cumulative){
    if (i < 0){
      return(0)
    } else {
      return(cumulative[20])
    }
  }
}

n <- 1000

#vector to store results of each simulation
game_outcome <- rep(0, n)

#for loop to run 100 simulations
for (i in 1:n){
  
  #play game and add to outcome
  game_outcome[i] <- game()
  
}

#expected outcome
expected <- mean(game_outcome)
print(expected)

plot(game_outcome)
```