---
title: "Simulate Monopoly Practice"
author: "Mikayla Mulgrew"
date: "2025-07-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Function is to make things different colours. (It allows for both LaTeX and html outputs). Use example: `r colorize("some words in red", "red")
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color,
      x)
  } else x
}
```

# Moving around the board

A Monopoly board has 40 spaces. Players take it in turns to roll two dice and traverse around the board according to the sum of the dice values. Simulate turns of a single player:

```{r}
#number of turns to be taken
turns <- 1000000

#initialize starting spot on board 0-GO
board_position <- 0

#create vectors to store die roll/size of move and board position numbers
move_sz <- rep(0, turns)
bd_positions <- rep(0, turns)

#for loop to simulate taking turns in monopoly

for (turn in 1:turns){
  
  #dice roll  using sample 
  roll <- sample(c(1:6), size=2, replace=TRUE)
  
  #sum die values 
  roll_value <- sum(roll)
  
  #add to current board position and update board position
  new_position <- board_position + roll_value
  #position is remainder when divided by 40 since there are 40 monopoly squares
  board_position <- (new_position %% 40)
  
  #store values in move_sz and bd_positions
  move_sz[turn] <- roll_value
  bd_positions[turn] <- board_position
}

#By increasing the number of turns taken, what distribution does the set of simulated board positions converge towards? Show this graphically using the histogram function. - Uniform Distribution

hist(bd_positions, breaks=seq(0,40,len=41), right=FALSE, main="Frequency of Board Positions", xlab = "Board Position")

```
# Going to Jail
#If a player lands on to Go To Jail space they must move immediately to the Jail space. Extend your code to include the possibility of going to jail. Here, assume that once in jail, the player continues as normal on the next turn.

```{r}

#number of turns to be taken
turns <- 1000000

#initialize starting spot on board 0-GO
board_position <- 0

#create jail positions 
go_to_jail <- 30
jail_position <- 10

#create vectors to store die roll/size of move and board position numbers
move_sz <- rep(0, turns)
bd_positions <- rep(0, turns)

#for loop to simulate taking turns in monopoly

for (turn in 1:turns){
  
  #dice roll  using sample 
  roll <- sample(c(1:6), size=2, replace=TRUE)
  
  #sum die values 
  roll_value <- sum(roll)
  
  #add to current board position and update board position
  new_position <- board_position + roll_value
  #position is remainder when divided by 40 since there are 40 monopoly squares
  board_position <- (new_position %% 40)
  
  #if player lands on go to jail, set board_position as jail_position
  if (board_position == go_to_jail){
    board_position <- jail_position
  }
  
  #store values in move_sz and bd_positions
  move_sz[turn] <- roll_value
  bd_positions[turn] <- board_position
}

#Now, what is the distribution of board positions during a long game?
hist(bd_positions, breaks=seq(0,40,len=41), right=FALSE, main="Frequency of Board Positions with Go To Jail", xlab = "Board Position")

```
## Questions

**Q1**. Can you explain the histogram of frequency against positions visited, qualitatively?
 
#After a large number of turns, 10,000 and above, the distribution of positions by frequency tends towards a uniform distribution. This means that is about equally likely to land on any of the board positions. 
  

**Q2**. You can also go to jail, if you roll three doubles (both dice having the same value) in a row. Update your code to allow for the possibility of going to Jail with three doubles. How does the distribution of board positions change?

#Distribution of board positions, with the exception of the go to jail and jail positions, still tend toward a uniform distribution since three doubles in a row isn't very common.

```{r}

#number of turns to be taken
turns <- 100000

#initialize starting spot on board 0-GO
board_position <- 0

#create jail positions 
go_to_jail <- 30
jail_position <- 10

#create vectors to store die roll/size of move and board position numbers
move_sz <- rep(0, turns)
bd_positions <- rep(0, turns)

#create another vector to hold boolean values for whether each given roll was a double or not, populate with FALSE values 
move_dbl <- rep(FALSE, turns)

#for loop to simulate taking turns in monopoly

for (turn in 1:turns){
  
  #dice roll  using sample 
  roll <- sample(c(1:6), size=2, replace=TRUE)
  
  #if die rolls are equal to each other than doubles for the move is changed to TRUE 
  if (roll[1] == roll[2]){
    move_dbl[turn] <- TRUE
  }
  
  #sum die values 
  roll_value <- sum(roll)
  
  #add to current board position and update board position
  new_position <- board_position + roll_value
  #position is remainder when divided by 40 since there are 40 monopoly squares
  board_position <- (new_position %% 40)
  
  #if player rolls three doubles in a row, send to jail_position
  #access last three role values, starting at turn >= 3
  if (turn >= 3){
    
    #get last three turns boolean values
    start <- turn - 1
    end <- turn - 3
    doubles <- move_dbl[start:end]
    
    #if all are TRUE, then send to jail_position
    if (all(doubles == TRUE)){
      board_position <- jail_position
    }
  }
  
  #if player lands on go to jail, set board_position as jail_position
  if (board_position == go_to_jail){
    board_position <- jail_position
  }  
  
#
  
  #store values in move_sz and bd_positions
  move_sz[turn] <- roll_value
  bd_positions[turn] <- board_position
}


#Now, what is the distribution of board positions during a long game?
hist(bd_positions, breaks=seq(0,40,len=41), right=FALSE, main="Frequency of Board Positions with Go To Jail and 3 Doubles To Jail", xlab = "Board Position")

```
