---
title: "Simulate Markov Chain Practice"
author: "Mikayla Mulgrew"
date: "2025-07-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("ggplot2")
```

#Define a vector called `x` which contains the numbers 1, 2, 3, and 4
#Use the function `sample` to pick from those four numbers in `x` 10,000 times
#Plot a histogram of the values

```{r}
#define a vector containing #s 1-4
x <- c(1,2,3,4)

#sample function to pick from those 4 numbers 10,000x, replacing values
samp_ <- sample(x, 10000, replace=TRUE)

#create a dataframe to be able to use ggplot 
df_2 <- data.frame(samp_, samp_)

#barplot of the frequency of each number picked
ggplot(df_2, aes(x=samp_)) + geom_bar()
```

#Use the `sample` or `sample.int` function to simulate values from rolls of an unbiased six-sided die. Show that the distribution of values you obtain is consistent with an unbiased die.

```{r}

#sample numbers 1-6, sides of the die, 10,000x
samp_int <- sample.int(6, 10000, replace = TRUE)

#create a data frame to be able to use in ggplot
samp_int_df <- data.frame(samp_int, samp_int)

#barplot for frequency of fair die rolls
ggplot(samp_int_df, aes(x=samp_int)) + geom_bar()

```

#If it is currently sunny, the probability of it raining the next day is 0.6. Conversely, if it is raining, the probability that it will become sunny the next day is 0.7 and 0.3 that it will continue raining.

```{r}

#transition matrix for probabilities of sun and rain
sunRain_Mat <- matrix(c(0.4, 0.7, 0.6, 0.3), nrow=2, ncol=2)

sunRain_Mat
```

#Create a simulation of 30 days of weather based on transtion matrix
# 1-sunny
# 2-rainy

#Can you extend this example to a three-state model?
# 1-sunny
# 2-rainy
# 3-cloudy

```{r}

#simulate 30 day weather probabilities based on transtion matrix
#initialise simulation with 1 - sunny
state <- 1

#create a vector of length 30 to hold weather state values
weather <- rep(0,30)

for (day in 1:30){
  
  #grab probabilities for transitions given current state
  trans_prob <- sunRain_Mat[state,]
  
  #use sample function to choose next state based on probability transitions
  state <- sample(c(1,2), size=1, prob=trans_prob)
  
  #store new weather state in vector
  weather[day]<-state
}

print(weather)


#tri-state model

#transition matrix to hold probability values for 3 weather states
tri_state_matrix = matrix(c(0.7, 0.2, 0.1,0.3, 0.3, 0.4,0.6, 0.2, 0.2), nrow=3, ncol=3, byrow=TRUE)

#initialise weather state at 1-sunny
i_state <- 1

#create a vector to hold weather state values for 30 days
tri_weather <- rep(0, 30)

for (day in 1:30){
  
  #grab probabilities for next weather state with current weather state
  tri_prob <- tri_state_matrix[i_state,]
  
  #use sample to choose next weather state using probabilities
  i_state <- sample(c(1,2,3), size=1, prob=tri_prob)
  
  #store new weather state in vector
  tri_weather[day] <- i_state
}

tri_weather
```