---
title: "U8L1 Two Sample Testing"
author: "Mikayla Mulgrew"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

Two methods of filling standard gas cylinders are claimed to be different.

In particular, process \(A\) is claimed to yield a higher pressure that process \(B\).

> A random sample of 72 cylinders were filled using process $A$, yielding $\bar{x}_A = 88$ and $s_A^2 = 4.5$.

> A random sample of 48 cylinders were filled with process $B$, yielding $\bar{x}_B = 79$ and $s_B^2 = 4.2$.

Conduct a hypothesis test, at the 5% significance level, to investigate this claim.

```{r}
#One Tail Test
#Ho: x_bar_A = x_bar_B
#Ha: x_bar_B < x_bar_A

#create given variables
x_bar_A <- 88
sd2_A <- 4.5
nA <- 72

x_bar_B <- 79
sd2_B <- 4.2
nB <- 48

alpha <- 0.05

#both sample sizes are large, pop. variance unknown - can apply CLT and use sample variances

#z-statistic
z <- (x_bar_A - x_bar_B)/sqrt(sd2_A/nA + sd2_B/nB)
print(z)

# 5% significance level
z_upper <- qnorm(1-alpha, mean=0, sd=1)
print(z_upper)

#Since z lies within the critical region, the null hypothesis is rejected. There is evidence to support that process A yields a higher pressure than process B.
```

# Problem 2

Two catalysts are available for a chemical process. Catalyst B is cheaper than catalyst A.

Provided catalyst B produces the same mean yields, it should be preferred.

To compare methods an experiment was conducted yielding the following data:

> \(A\): 91.50 94.18 92.18 95.39 91.79 89.07 94.72 89.21

> \(B\): 89.19 90.95 90.46 93.21 97.19 97.04 91.07 92.75

Is there evidence to say the two catalysts produce different mean yields? Test at the 5% significance level
```{r}
#creating dataset A and its variables
data_A <- c(91.50, 94.18, 92.18, 95.39, 91.79, 89.07, 94.72, 89.21)
xbarA <- mean(data_A)
sdA <- var(data_A)
nA <- length(data_A)

#creating dataset B and its variables
data_B <- c(89.19, 90.95, 90.46, 93.21, 97.19, 97.04, 91.07, 92.75)
xbarB <- mean(data_B)
sdB <- var(data_B)
nB <- length(data_B)

#Ho: xbarA = xbarB
#Ha: xbarA != xbarB

#will assume normality, small sample size so using pooled estimator of variance, will also assume difference of sample means is same as difference in population means

sp2 <- ((nA-1)*sdA + (nB-1)*sdB)/(nA + nB - 2)

#t-statistic 
t <- (xbarA-xbarB)/sqrt(sp2*((1/nA) + (1/nB)))
print(t)

one_tail <- qt(1-0.025, df=nA+nB-2)
print(one_tail)

#Sine t does not fall within the rejection region, there is no reason to reject the null hypothesis that the mean yields of catalyst A and B are different. 

```